Box 1 - TDT Curves

### 2023-09-25 ms/ms.qmd
First draft of box

We can have a look at each new species' TDT curves collected by Vanessa (@fig-tdt). We can also use the data to calculate the TDT curve (slope) and $CT_{max}$ for each species (@tbl-1).

@@

```{r, correlations}

# Estimate phylogenetic correlatons between traits. 
    # Trait variances 
   phylocor_ctmax_tdt <- phylo_corrs(vcv, trait1 = "tdt_z", trait2 = "ctmax")
   posterior.mode(phylocor_ctmax_tdt)
   HPDinterval(phylocor_ctmax_tdt)

@@

We integrated static lethal limits ($CT_{max}$) collected by @Heerwaarden2021 and XX with dynamic lethal limit measures from TDT curces [i.e., slope ($z^{TDT}$) and intercept ($CT_{max}^{TDT}$)] estimated by XX. We also collected data to estimate TDT curve parameters for eight additonal species following protocols outlined in @Orsted2022. Critical fertility limits for *Drosophilids* were compiled from @Parratt2021 and @Heerwaarden2021. These data were integrated with data on lethal limits at the species-level. Overall, we obtained data for `r length(unique(t$tip.label))` species ($N_{CT_{max}}$ = `r num_spp[2]-5` species; $N_{z^{TDT}}$ = `r num_spp[4]-4` species and $N_{CFLs}$ = `r num_spp[3]-4` species). When the same species were sampled multiple times we retained within species variation in models. We then fit the following model in R (vers. `r paste0(R.Version()$major, ".", R.Version()$minor)`)[@R2023] with the package *MCMCglmm* (vers. `r utils::packageVersion("MCMCglmm")`), which uses data augmentation during model fittng to estimate missing data [@Hadfield2010]:

@@

$$
\begin{aligned}
& p_{k} \sim N\left(0, \sigma_{p}^2 \textbf{A} \right) \\
& sp_{k} \sim N\left(0, \sigma_{sp}^2\textbf{I} \right) \\
& e_{i,k} \sim N\left(0,  \sigma_{e}^2\textbf{I}\right) \\
\end{aligned}
$$

where $p_{k}$ and $sp_{k}$ are phylogenetic (evolutionary) and population level effects (assumed to be sampled from a normal distribution, $\sim N\left(0, \sigma_{sp/p}^2 \right)$) with the evolutionary relatedness being structured according to the phylogenetic relatesness matrix ($\textbf{A}$).

>>>>

We can have a look at each new species' TDT curves collected by Vanessa (@fig-tdt). We can also use the data to calculate the TDT curve (slope) and $CT_{max}$ for each species (Table 1).

@@

```{r, correlations}
#| label: correlations
#| echo: false
#| include: false
# Estimate phylogenetic correlatons between traits. 
    # Trait variances 
   phylocor_ctmax_tdt <- phylo_corrs(vcv, trait1 = "tdt_z", trait2 = "ctmax")
   posterior.mode(phylocor_ctmax_tdt)
   HPDinterval(phylocor_ctmax_tdt)

@@

We integrated static lethal limits ($CT_{max}$) collected by @Heerwaarden2021 and XX with dynamic lethal limit measures from TDT curces [i.e., slope ($z^{TDT}$) and intercept ($CT_{max}^{TDT}$)] estimated by XX. We also collected data to estimate TDT curve parameters for eight additonal species following protocols outlined in @Orsted2022. Critical fertility limits for *Drosophilids* were compiled from @Parratt2021 and @Heerwaarden2021. These data were integrated with data on lethal limits at the species-level. Overall, we obtained data for `r num_spp[1]` species ($N_{CT_{max}}$ = `r num_spp[2]` species; $N_{z^{TDT}}$ = `r num_spp[4]` species and $N_{CFLs}$ = `r num_spp[3]` species). When the same species were sampled multiple times we retained within species variation in models. We then fit the following model in R (vers. `r paste0(R.Version()$major, ".", R.Version()$minor)`)[@R2023] with the package *MCMCglmm* (vers. `r utils::packageVersion("MCMCglmm")`), which uses data augmentation during model fittng to estimate missing data [@Hadfield2010]:

@@

$$
\begin{aligned}
 & p_{k} \sim N\left(0, \sigma^2_{p} \textbf{A}\right) \\ 
 & sp_{k} \sim N\left(0, \sigma^2_{sp} \textbf{I}\right) \\
 & e_{i,k} \sim N\left(0, \sigma^2_{e} \textbf{I}\right)
\end{aligned}
$$

where $u_{i,k}$ is the overall mean (intercept) for each trait for each observation, *i*, for species, *k*; $p_{k}$ and $sp_{k}$ are phylogenetic (evolutionary) and species effects (i.e., ecological species effects) (assumed to be sampled from a normal distribution) with the evolutionary relatedness being structured according to the phylogenetic relatedness matrix ($\textbf{A}$) as opposed to an identity matrix ($\textbf{I}$). Heat tolerance measures are modelled according to a multivariate normal distribution. Covariances between thermal tolerance measures at each random effect level are also estimated.
